---
---
---
title: "Photosynthesis lab"
author: "Senanu Spring-Pearson"
date: "`r Sys.Date()`"
output: html_document
parent: R
grand_parent: Labs
---

```{r setup, include=TRUE, echo=FALSE}
knitr::opts_knit$set(global.device = TRUE, echo=TRUE) # leave this alone
```

As you have done before, please make sure you can successfully knit this document before you make any changes to it.

Take out my name in the headers and put your name in it's place.

## Part 1: Effects of climate change on photosynthesis

For this part, you do not need to do any real data analysis.

1. Briefly describe your experiment. What variable did you test? How did you change that variable?

    Replace this text with your answer. If you insert 4 spaces at the beginning, it'll be nicely formatted.

2. Describe how the 4 beakers differed:

    Control:

    Beaker 2:

    Beaker 3:
    
    Beaker 4:

3. What were the photosynthesis rates?

    Control:
    Beaker 2:
    Beaker 3:
    Beaker 4:

4. Which had the fastest and slowest rates of photosynthesis?

    Replace this text with your answer. If you insert 4 spaces at the beginning, it'll be nicely formatted.

5. What can you conclude about how climate change will affect photosynthesis?

## Part 3: Absorption spectrum of elodea

Make a graph showing the absorption spectrum of the pigments in elodea. You should:

1. Save a google sheets data table with the headers "wavelength" and "absorbance". Remember not to include extraneous characters like parentheses or spaces.
2. Read the data into this document and save it in a variable called `dat.spec` (for 'spectrum'). It is important to use a different name than what you used in part 1 above.
3. Use the `plot()` command to generate a figure showing how the absorbance varies with the wavelength.
    1. remember to tell R which data column you want on which axis. You can do this with the `x=dat$variable1` and `y=dat$variable2` arguments to the `plot()` command.
    2. Add axis labels with `xlab="my X-axis label"` and similarly for `ylab=`.
    3. Connect the points with `type='l'` (that's the lower-case letter 'l' which stands for 'lines'). As with the `xlab` and `ylab` arguments, this should be an argument to the `plot()` command.
4. When the code chunk below works, change "eval=FALSE" to "eval=TRUE"
5. Knit this document and turn it in.

```{r absorbancePlot, eval=FALSE, echo=TRUE, include=TRUE}
dat <-
plot()
```

## What to turn in...

Knit this document. ** Make sure that the final version has 1 graph**. If you don't have a graph, it may be because you didn't change `eval=FALSE` to `eval=TRUE`, so check that those are all changed.
